Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-11-27 dsn-2018/bare_conf.tex
Updated abstract and title
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Measuring the Utility of a Collision:\\ The Neighborhood of MD4 Collisions}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
We propose a series of metrics for evaluating the utility of collision classes
and develop a set of techniques for generating new collision classes in hash
functions utilizing a logical cryptanalaysis framework. We demonstrate the
effectiveness of this framework on the hash function MD4 and show that the
existing attacks are of high utility, but do not allow for second preimage
attacks under an expansive neighborhood. We partially analyze over 35,000 new
collision classes for their utility under these methods.
\end{abstract}

@@

@@

\subsection{Distance Metric}

We introduce a distance function between collision classes by the number of
differences in intermediate rounds deltas. We find justification for this
metric in the existing literature on MD4: H. Dobbertin's \cite{Dobbertin1998}
and P. Kasselman's \cite{KasselmanMD4} classes have distance $1$ under this
metric.

That is:
\begin{align}
    \delta(C_1, C_2) = \| \{ i : I(C_1, i) \neq I(C_2, i) \} \|
\end{align}

For the example collisions, below are distances under this metric:
\textbf{TODO}.

\subsection{Neighborhood}

We define the neighborhood of a collision class to be all other collisions
which have a distance of one.

That is:
\begin{align}
    N(C) = \{ C_j : \delta(C, C_j) = 1 \}
\end{align}

For the example collisions, below are sizes of the neighborhoods:
\textbf{TODO}.

@@

\textbf{TODO} - Describe technique applied to tags

@@

@@

Include graphic of count of unique differentials for all 35k collision classes

Separate table for known attacks as data points. Examples in appendix?

\subsection{Unit-Step Neighborhood}

Include graphic of count of number of neighbors for all 35k collision classes

Separate table for known attacks as data points.

@@

>>>>

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The discovery of fast collision attacks in cryptographic hash functions has
traditionally resulted in the immediate deprecation of that hash function.
In this paper we propose five scalable and practical metrics for evaluating
the utility of collision classes based on boolean constraints and show that
the published attacks by X. Wang, Y. Sasaki, P. Kasselman, H. Dobbertin, and
M. Schlaffer in MD4 have high utility. We expand on existing attacks by
developing a series of techniques based on logical cryptanalysis to find
over 35,000 collisions in MD4 based on existing collisions, through the novel
definition of a collision neighborhood. We demonstrate new techniques for
inductively building full collisions from reduced round variants of MD4. We
propose these techniques as a mechanism for measuring hash trustworthiness and
discuss potential applications to real-world systems.
\end{abstract}

@@

$\mathcal{C}$ for collision class. Ordered tuple.
$\mathcal{F}$ for family of collision classes. Set.
$I$ is an indexing function.

@@

\subsection{Distance Metrics}

We introduce a distance function, $\delta$ between collision classes by the
number of differences in intermediate rounds deltas. That is, given two
collision classes, $C_1, C_2 \in \mathcal{C}$:
\begin{align}
    \delta(C_1, C_2) = \big| \{ i : I(C_1, i) \neq I(C_2, i) \} \big|
\end{align}
We find justification for this metric in the existing literature on MD4:
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes have distance $1$ under this metric.

Please refer to Table \ref{table:distance} for the distances between existing
collisions in MD4.

\begin{table}
    \caption{Distance between Existing Collision Classes}
    \label{table:distance}
    \begin{tabular}{c c c c c c}
        & X. W. & Y. S. & P. K. & H. D. & M. S. \\
        X. Wang's & 0 & 21 & 26 & 27 & 12 \\
        Y. Sasaki's & & 0 & 25 & 26 & 2 \\
        P. Kasselman's &  &  & 0 & 1 & 25 \\
        H. Dobbertin's &  &  &  & 0 & 26 \\
        M. Schlaffer's &  &  &  &  & 0 \\
    \end{tabular}
\end{table}

We can define a similar distance function, $\Delta$, between families of
collision classes by cardinality of the symmetric difference in the two
collision families. That is, given two families of collision classes,
$F_1, F_2 \in \mathcal{F}$:
\begin{align}
    \Delta(F_1, F_2) = \big| (F_1 \cup F_2) \setminus (F_1 \cap F_2) \big|
\end{align}
This is convenient for when the specifics of the differential path do not
matter, merely that there exist at least one collision class of the specified
form.

\subsection{Neighborhoods}

We define the neighborhood of a collision class, $C \in \mathcal{C}$, to be the
set of all other collisions at a fixed distance, $d \in \mathbb{N}$, from $C$.
That is:
\begin{align}
    N(C, d) & = \{ C_j : \delta(C, C_j) = d \} \\
    N(C) & = N(C, 1)
\end{align}
For instance, $C_{Wang} \in N(C_{Sasaki}, 21)$. The distance parameter, $d$,
may optionally be ommited, in which case the unit distance is implied. Thus,
$C_{Kasselman} \in N(C_{Dobbertin})$.

The implications of neighborhoods are discussed further in section
\ref{empirical:neighborhood}.

We can similarly define the neighborhood of a family of collision classes,
$F \in \mathcal{F}$, to be the set of all other families of collision classes
at a fixed distance, $d \in \mathbb{N}$, from $F$.

@@

\textbf{TODO} - Describe technique applied to tags.

@@

\textbf{TODO}
Please refer to Table \ref{table:differentials} for the number of differential
paths...

\begin{table}
    \caption{Number of Differentials for Existing Collision Classes}
    \label{table:differentials}
    \begin{tabular}{c c c c}
        \textbf{Attack} & \textbf{Size} & \textbf{Attack} & \textbf{Size} \\
        X. Wang's & 64 & Y. Sasaki's & 4 \\
        H. Dobbertin's & 32 & P. Kasselman's & 32 \\
        M. Schlaffer's & 64 & & \\
    \end{tabular}
\end{table}

\subsection{Unit-Step Neighborhood} \label{empirical:neighborhood}

@@

Please refer to Table \ref{table:neighborhood} for the sizes of neighborhoods
of existing collisions. In particular, note that Y. Sasaki's attack--which
claimed to improve upon X. Wang's attack--has a larger neighborhood size, and
likewise with P. Kasselman's attack which improved upon H. Dobbertin's attack.

\begin{table}
    \caption{Neighborhood Sizes for Existing Collision Classes}
    \label{table:neighborhood}
    \begin{tabular}{c c c c}
        \textbf{Attack} & \textbf{Size} & \textbf{Attack} & \textbf{Size} \\
        X. Wang's & 54 & Y. Sasaki's & 157 \\
        H. Dobbertin's & 55 & P. Kasselman's & 60 \\
        M. Schlaffer's & 100 & & \\
    \end{tabular}
\end{table}

@@

See Table \ref{table:ascii} for results. Note that the validation of
P. Kasselman's and H. Dobbertin's results do not hold for the latter attacks
by X. Wang, Y. Sasaki, or M. Schlaffer. Thus, while the latter attacks have
been viewed as being of better quality, under this particular metric,
P. Kasselman's and H. Dobbertin's collision classes are better.

\begin{table}
    \caption{ASCII Blocks in Existing Collision Classes}
    \label{table:ascii}
    \begin{tabular}{c c c}
        \textbf{Attack} & \textbf{Single Block} & \textbf{Both Blocks} \\
        X. Wang's & true & false \\
        Y. Sasaki's & true & false \\
        P. Kasselman's & true & \textbf{true} \\
        H. Dobbertin's & true & \textbf{true} \\
        M. Schlaffer's & true & false \\
    \end{tabular}
\end{table}
